Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-06-01 doc/paper.tex
added some references and started Potential Benefits

As early as Jelinek's (in)famous quote during the 1980's that ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing. 
More recently, for example, \mynewcite{collobert:2011} mention the possibility of ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even the most fundamental of linguistic concepts.
A specific instance of this overall trend is the gradual shift of machine translation (MT) away from classical linguistic rule-based approaches \mycitep{hutchins:2005} to contemporary machine learning approaches, which underlie statistical phrase-based systems like Moses \mycitep{moses} and more recent neural network systems \mycitep{sutskever:2014,bengio:2014}.  
But in the end, even the most cutting-edge machine learning MT system relies crucially on multilingual parallel corpora, typically taken from external sources like multilingual parliamentary proceedings. 
This results in what we perceive to be a disproportionate amount of effort being dedicated to the machine learning component of machine translation, with input corpora being treated almost as an afterthought.\footnote{
    Of course, these parallel corpora represent hundreds of translator-hours of work. 
    However, these translations are generally optimized for human consumption, not machine learning, with no regard for such niceties as sentence alignment.
    This is illustrated briefly in Section \ref{case_study}.
    }

@@

% this is some stupid little side note I just noticed a couple weeks ago, but it is expanding to a huge size, because it's like the only concrete ``positive'' result I have...
\subsection{Case Study}
\label{case_study}

@@

As an example of suboptimal parallel corpora being used for machine translation, consider the official Chinese to English baseline translation of the {\small \tt tst2010} data set\footnote{https://wit3.fbk.eu/score.php?release=2014-01} from the TED talks corpus, described in greater detail in Section \ref{experiments}.
As a very crude way of assessing translation quality, we used the official evaluation software MultEval \mycitep{multeval} to compute scores for each translated line individually.
Upon manual inspection, the worst scoring individual lines (with TER scores over 150\%) were almost all found to correspond to poor human translations on the Chinese side of the gold corpus.\footnote{
    High TER scores indicated misaligned sentences as well as clerical errors.  In particular, the lecture with {\small \tt talkid} 767 had dozens of erroneous lines, in some of the worst cases having the same Chinese sentence copy/pasted three times on the same line. 
    The baseline Moses system tries valiantly, but it ultimately scores poorly for trying to produce a good machine translation of a poor human translation.
    Alternatively, one can interpret these errors as an additional source of noise in the human encoding of English into Chinese.
}
Such translations might suffice for viewing translated videos monolingually, but they are far from ideal for training a machine translation system.

@@

\subsection{Upsides, Potential Applications - RETITLE}

Okay, so Higher corpus quality is desirable.
Larger in-domain corpus quantity is desirable too - see learning curve controls below?

My main point seems to be that nobody makes high quality and quantity parallel corpora JUST for machine translation, and it seems like that's sorely needed for both training and benchmarking.

Traditionally, corpus construction has been considered prohibitively expensive, ``scavenged'' from other sources like parliamentary proceedings or lecture subtitles.
But we argue here that it need not be so expensive.

Such synthetic parallel corpora can also be viewed as benchmark tasks (need to cite \mycitep{Facebook bAbI}).
This is actually probably more important than you'd think, because looking at IWSLT15, some of the human translations are VERY loose.
So there's another element of noise in the data, and it's one that no amount of downstream machine learning can get rid of.
(Although people have tried corpus evaluation...)

% interesting, but probably confusing to readers?
%In the language of agile software development, test-driven MT system development, in which linguists pose sentence pairs, and MT developers work to translate them

This could take the form of amplified training sets in the short run, and perhaps a ``universal translation memory'' in the long run.

methods every few years; linguistic theories, representations every decade or so; parallel corpora live forever

@@

unlike all the different competing methods, it seems that a ``universal translation memory'' would be inherently collaborative (methods are competitive by nature, whereas data can be collaborative)

@@

@@

The idea underlying generation system was essentially proposed 
This aspect of our work is perhaps closest in spirit to the equivalence classes used by \mynewcite{brown:99} to reduce the data requirements of an EBMT system \mycitep{ebmtbook}.

@@

other approaches to synthetic corpora, which are essentially bootstrapped from other MT systems

@@

\section{Our Primitive Multilingual Generator}

@@

%\subsection{Implementation Details}
Python
YAML for data: very well suited. full Unicode support, and you don't even need to type quotation marks to enclose strings
Crude though it may be, we plan on eventually open-sourcing both the generator and its accompanying data

\section{Preliminary Experiments}
\label{experiments}

\subsection{Methods}
As mentioned in 
The IWSLT 2015 version of the WIT$^3$ corpus, which consists of multilingual subtitles of TED talks, covering a wide variety of subjects. 
considered briefly in Section \ref{case_study}

@@

\section{Proposed Future Work}

@@

@@

But must this be so? Vast untapped resource of students or whatever.

>>>>

As early as Jelinek's (in)famous quote during the 1980's that ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing. 
More recently, for example, \mynewcite{collobert:2011} mention the possibility of ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even the most fundamental of linguistic concepts.
A specific instance of this overall trend is the gradual shift of machine translation (MT) away from classical linguistic rule-based approaches \mycitep{hutchins:2005} to contemporary machine learning approaches, which underlie statistical phrase-based systems like Moses \mycitep{moses} and more recent neural network systems \mycitep{sutskever:2014,bengio:2014}.  
But in the end, even the most cutting-edge machine learning MT system relies crucially on multilingual parallel corpora, typically taken from external sources like multilingual parliamentary proceedings. 
This results in what we perceive to be a disproportionate amount of effort being dedicated to the machine learning component of machine translation, with input corpora being treated almost as an afterthought.\footnote{
    Of course, these parallel corpora represent hundreds of translator-hours of work. 
    However, these translations are generally optimized for human consumption, not machine learning, with no regard for such niceties as sentence alignment.
    This is illustrated briefly in Section \ref{subsec:case_study}.
    }

@@

% although it is ``concrete'', this section will probably be the first to get cut if I run OVER length (which had previously been unimaginable, but is starting to look more and more likely)
% this is some stupid little side note I just noticed a couple weeks ago, but it is expanding to a huge size, because it's like the only concrete ``positive'' result I have...
\subsection{Case Study}
\label{subsec:case_study}

@@

As an example of suboptimal parallel corpora being used for machine translation, consider the official Chinese to English baseline translation of the {\small \tt tst2010} data set\footnote{https://wit3.fbk.eu/score.php?release=2014-01} from the TED talks corpus, described in greater detail in Section \ref{sec:experiments}.
As a very crude way of assessing translation quality, we used the official evaluation software MultEval \mycitep{multeval} to compute scores for each translated line individually.
Upon manual inspection, the worst scoring individual lines (with TER scores over 150\%) were almost all found to correspond to poor human translations on the Chinese side of the gold corpus.\footnote{
    High TER scores indicated misaligned sentences as well as clerical errors.  In particular, the lecture with {\small \tt talkid} 767 had dozens of erroneous lines, in some of the worst cases having the same Chinese sentence copy/pasted three times on the same line. 
    The baseline Moses system tries valiantly, but it ultimately scores poorly for trying to produce a good machine translation of a poor human translation.
    Alternatively, one can interpret these errors as an additional source of noise in the human encoding of English into Chinese.
}
Such translations might suffice for viewing translated videos monolingually, but they are far from ideal for training a machine translation system.

@@

\subsection{Potential Benefits}

Large, high quality synthetic parallel corpora could be useful as:

\begin{itemize}
\item Training sets for corpus-based MT systems, particularly for low-resource language pairs
\item Benchmark tasks for MT evaluation, in the spirit of the Facebook bAbI tasks in question answering \mycitep{bAbI}
\item Translation memories for computer-aided translation
\item ``Language archives'': New NLP methods are developed every few years; linguistic theories change every decade or two; parallel corpora live forever (e.g., the Rosetta Stone).
\end{itemize}
% sigh, this set of bullets seems much clearer than all those stupid sentences that came before

Larger in-domain corpus quantity is desirable too - see learning curve controls below?

@@

in the spirit of example-based MT system \mycitep{ebmtbook}

@@

% should this go in the ``conclusion''?
Okay, so our ultimate goal is translation of everyday conversational sentences

This could take the form of amplified training sets in the short run, and perhaps a ``universal translation memory'' in the long run.
unlike all the different competing methods, it seems that a ``universal translation memory'' would be inherently collaborative (methods are competitive by nature, can't combine naturally, outside of ensemble methods, whereas data provides a natural avenue for large-scale collaboration)

Section \ref{sec:related} discusses previous related work,
Section \ref{sec:generator} describes our first attempt at implementing a parallel corpus generator,
Section \ref{sec:experiments} gives preliminary results, 
and Section \ref{sec:future} proposes future work.

@@

The idea underlying generation system was essentially proposed

@@

Previous attempts at creating synthetic parallel corpora have used rule-based MT (RBMT) systems \mycitep{hu:2007,dugast:2008,murakami:2009,rubino:2014}

but the input to the RBMT systems was again dictated by external corpora -- domain overlap was prioritized over linguistic consideration of what sentences the RBMT system could translate most accurately.

Our approach,

not designed from the ground up. We consider a more {\em ab initio} approach

This aspect of our work is perhaps closest in spirit to the equivalence classes used by \mynewcite{brown:99} to reduce the data requirements of an EBMT system \mycitep{ebmtbook}.

But we propose allowing more rule-based modifications of these templates
Also, our goal is to generate a large, wide-coverage, high-quality corpus, not jumping directly to translation from examples as in EBMT systems

@@

\section{A Primitive Multilingual Generator}
\label{generator}

@@

%\subsection{Implementation Details}
Python
YAML for data: very well suited. full Unicode support, and you don't even need to type quotation marks to enclose strings.
with an eye toward scalability
Crude though it may be, we plan on eventually open-sourcing both the generator and its accompanying data

\section{Preliminary Experiments}
\label{sec:experiments}

\subsection{Methods}
As mentioned in 
The IWSLT 2015 version of the WIT$^3$ corpus, which consists of multilingual subtitles of TED talks, covering a wide variety of subjects. 
considered briefly in Section \ref{subsec:case_study}

@@

\section{Proposed Future Work}
\label{sec:future}

@@

parallel corpus construction has been considered prohibitively expensive, hence the use of pre-translated sources like parliamentary proceedings
we believe it is worth an initial investment to create a scalable system

@@

But must this be so? Vast untapped resource of students or whatever.
an open challenge to the NLP community at large
with a sufficiently compelling and thus widely used educational app 
a potentially game-changing source of supervised data for NLP in general
